For Finitary Induction-Induction, \\ Induction is Enough

### 2020-05-01 paper/paper.tex
stuff

  $\Con$, $\Ty$, $\Sub$, and so on, one per row. Each such components
  has three parts denoted by $^\A$, $^\D$ and $^\S$. The equality
  components of $\SignAlg$ are omitted as they are all reflexivity.

@@

@@

\begin{definition}\label{def:theoryofsignatures}
  A model of ETT supports the theory of IIT signatures if there is a
  signature algebra $\I : \SignAlg$ equipped with a unique morphism
  into any algebra $\rec : (M:\SignAlg)\ra\SignMor\,\I\,M$. We call
  this algebra the syntax or initial algebra, the morphism is called
  recursor.
\end{definition}

>>>>

  $\Con$, $\Ty$, $\Sub$, and so on, one per row. Each such component
  has three parts denoted by $^\A$, $^\D$ and $^\S$, respectively. The
  equality components of $\SignAlg$ are omitted as they are all
  reflexivity.

@@

\begin{definition}[The set signature algebra $\A$]\label{def:a}
  $\A : \SignAlg$ is given by the first $^\A$ components of $\ADS$
  (\Cref{def:ads}), that is, $\Con^\A := \Set$,
  $\Ty^\A\,\Gamma := \Gamma\ra\Set$,
  $\Sub^\A\,\Gamma\,\Delta := \Gamma\ra\Delta$, and so on. Note that
  there is a morphism from $\ADS$ to $\A$ defined by $\blank^\A$ at
  each component, which we also denote by
  $\blank^\A : \SignMor\,\ADS\,\A$.
\end{definition}

@@

\begin{definition}\label{def:theoryofsignatures}
  A model of ETT supports the theory of IIT signatures if there is a
  signature algebra $\I : \SignAlg$ equipped with a unique morphism
  $\ll\blank\rr_{M} : \SignMor\,\I\,M$ into any algebra $M$. Sometimes
  we omit the subscript $_M$. We call $\I$ the syntax or initial
  algebra, the morphism $\ll\blank\rr$ is called recursor.
\end{definition}

First we define the term signature algebra using which we obtain the
constructors $\con$ for any IIT signature. Then we will define another
signature algebra which provides the eliminator. Before doing these,
we illustrate the idea on natural numbers.
\begin{example}
  For natural numbers, we will define the constructors $\con$ as the
  following natural number algebra
  $(\mathsf{Nat},\mathsf{zero},\mathsf{suc})$. We write variable names
  instead of de Bruijn indices for readability.
  \begin{alignat*}{5}
    & \mathsf{Nat} && := \Tm^\I\,(\cdot\rhd N:\U\rhd z:\El\,N\rhd s:N\Ra\El\,N)\,(\El\,N) \\
    & \mathsf{zero} && := z \\
    & \mathsf{suc} && := \lambda t. (s\oldapp t)
  \end{alignat*}
  Natural numbers are simply $\I$-terms of type $\El\,N$ in the
  context representing the signature for natural numbers. In this
  context, the only way to define a term of type $\El\,N$ is to use
  $z$ and $s$, corresponding to the $\mathsf{zero}$ and $\mathsf{suc}$
  constructors.

  To define the action of the eliminator on a natural number
  $n : \mathsf{Nat}$, let's look at the type of the displayed algebra
  interpretation of the number:
  \[
    {\ll n \rr_\ADS}^\D : (\gamma^D : \ll\cdot\rhd N:\U\rhd z:\El\,N\rhd s:N\Ra\El\,N\rr^\D\,\con) \ra \ll\El\,N\rr^\D\,(\ll n\rr^\A\,\con)
  \]
  This says that for a displayed algebra $(N^D,z^D,s^D)$ over $\con$
  (i.e.\ the motives and methods of the eliminator), we get a witness
  of the predicate $\ll\El\,N\rr^\D = N^D$ at the algebra
  interpretation of $n$. This is not yet good, as we would like to get
  $N^D\,n$ instead of $N^D\,(\ll n\rr^\A\,\con)$ as a result. However,
  interpretation into the term signature algebra will imply that
  $n = \ll n\rr^\A\,\con$.
\end{example}

\begin{definition}[Term signature algebra $\A\blank\C_{\blank}$]
  For an $M:\SignAlg$ and $\Omega:\Con^M$, we define ${\A M\C}_\Omega$
  which we call the term signature algebra. It is equipped with
  morphisms $\blank^\A : \SignMor\,({\A M\C}_\Omega)\,\A$ and
  $\blank^M : \SignMor\,({\A M\C}_\Omega)\,M$.

  We define ${\A M\C}_\Omega$ by listing all its components $\Con$,
  $\Ty$, $\Sub$, and so on, one per row. Each component has three
  parts denoted by $^\A$, $^M$ and $^\C$. The $^\A$ and $^M$ parts are
  just reusing the corresponding parts from $\A$ and $M$,
  respectively, and thus the morphisms $\blank^\A$ and $\blank^M$ are
  defined as the obvious projections. We omit the equality components,
  as they come from UIP or are trivial. We also omit the components
  for terms and substitutions as their $^\C$ parts are uninformative
  equational reasonings.
  \begin{alignat*}{10}
    & \Gamma^\A:\Set && \hspace{0.4em} && \Gamma^M:\Con^M && \hspace{0.4em} && \Gamma^\C: \Sub^M\,\Omega\,\Gamma^M\ra \Gamma^\A \\
    & A^\A:\Gamma^\A\ra\Set && && A^M:\Ty^M\,\Gamma^M && && A^\C:(\nu:\Sub^M\,\Omega\,\Gamma^M)\ra \\
    & && && && && \hspace{1em} \Tm^M\,\Omega\,(A^M[\nu])\ra A^\A\,(\Gamma^\C\,\nu) \\
    & \sigma^\A:\Gamma^\A\ra\Delta^\A && && \sigma^M:\Sub^M\,\Gamma^M\,\Delta^M && && \sigma^\C:\Delta^\C\,(\sigma^M\circ\nu)=\sigma^A\,(\Gamma^\C\,\nu) \\
    & t^\A:(\gamma:\Gamma^\A)\ra A^\A\,\gamma && && t^M:\Tm^M\,\Gamma^M\,A^M && && t^\C : A^\C\,\nu\,(t^M[\nu])= t^A\,(\Gamma^\C\,\nu) \\
    & (A[\sigma])^\A\,\gamma := A^\A\,(\sigma^\A\,\gamma) && && (A[\sigma])^M := A^M[\sigma^M]^M && && (A[\sigma])^\C\,\nu\,t := A^\C\,(\sigma^M\circ\nu)\,t \\
    & \cdot^\A := \top && && \cdot^M := \cdot^M && && \cdot^\C\,\nu := \tt \\
    & (\Gamma\rhd A)^\A := && && (\Gamma\rhd A)^M :=  && && (\Gamma\rhd A)^\C\,\nu := \\
    & \hspace{1em} (\gamma:\Gamma^\A)\times A^\A\,\gamma && && \hspace{1em} \Gamma^M\rhd^M A^M && && \hspace{1em} (\Gamma^\C\,(\pi_1\,\nu), A^\C\,(\pi_1\,\nu)\,(\pi_2\,\nu)) \\
    & \U^\A\,\gamma := \Set && && \U^M := \U^M && && \U^\C\nu\,a := \Tm^M\,\Omega\,(\El^M\,a) \\
    & (\El\,a)^\A\,\gamma := a^\A\,\gamma && && (\El\,a)^M := \El^M\,a^M && && (\El\,a)^\C\,\nu\,t := (a^\C\,\nu) \# t \\
    & (\Pi\,a\,B)^\A\,\gamma := && && (\Pi\,a\,B)^M := && && (\Pi\,a\,B)^\C\,\nu\,t :=  \\
    & \hspace{1em} (\alpha:a^\A\,\gamma)\ra B^\A\,(\gamma,\alpha) && && \hspace{1em} \Pi^M\,a^M\,B^M && && \hspace{1em} \lambda\alpha.B^\C\,\big(\nu, (a^\C\,\nu)\#\alpha\big)\,\big(t\oldapp ((a^\C\,\nu) \#\alpha)\big) \\
    & (\Pim\,T\,B)^\A\,\gamma := && && (\Pim\,T\,B)^M := && && (\Pim\,T\,B)^\C\,\nu\,t := \lambda\alpha.(B\,\alpha)^\C\,\nu\,(t\appm\alpha) \\
    & \hspace{1em} (\alpha:T)\ra(B\,\alpha)^\A\,\gamma && && \hspace{1em} \Pim^M\,T\,B^M && && 
  \end{alignat*}
\end{definition}

xxx
